Meta analysis of reprocessed RNASeq data from AMP-AD for hg38 Reprocessed Data (all 7 brain regions)

### 2021-11-17 code/RNA-Seq_Meta_Analysis.Rmd
New version RNA-Seq Weights

```{r synapse.parameters, include=FALSE, cache=TRUE}
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

#parentId = 'syn18936948';
parentId = 'syn22351719';
activityName = 'Meta analysis of differential expression hg38 Reprocessed RNASeq';
activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data hg38 Reprocessed Data (6 brain regions - No CBE)';
thisFileName <- 'RNA-Seq_Meta_Analysis.Rmd'
# Github link
thisRepo <- getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

@@

```{r get.exp}
# Function to read tsv/csv files from synapse
downloadFile <- function(id){ 
  fread(syn_temp$get(id)$path, data.table = F)
}
# Read covariates
covar.ids = c(ROSMAP = 'syn8456631', MSSM = 'syn8484996', MAYO = 'syn8466814')
all.used.ids = as.character(covar.ids)
covar =  covar.ids %>%
  lapply(downloadFile)

# Read adjusted expression
expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
all.used.ids = c(all.used.ids, as.character(expr.ids))
expr =  expr.ids %>%
  lapply(downloadFile) %>%
  plyr::join_all(type = 'inner')
```

```{r modify.covar}
covar$ROSMAP = covar$ROSMAP %>%
  dplyr::select(SampleID, Sex, Diagnosis, APOE4) %>%
  dplyr::mutate(Tissue = 'DLPFC')
covar$MSSM = covar$MSSM %>%
  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
  dplyr::rename(Sex = Sex) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
covar$MAYO = covar$MAYO %>%
  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
covar = covar %>%
  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study') %>%
  dplyr::select(-Tissue1)
```

## Perform fixed and random effect meta-analysis for each gene across 7 different brain regions
#### Meta-analysis of AD-CONTROL
```{r ad.cntrl}
# Mean and sd of gene expression in each tissue x diagnosis group
tissue.dx.summary = plyr::ddply(covar, .(Tissue, Diagnosis), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison - Exclude CBE
meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary[(tissue.dx.summary$Tissue%in%'CBE')==F,], .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$Tissue
  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
  rownames(cntrl.effect) = cntrl.effect$Tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

@@

#### Adjust the AD-Control Model for weights
```{r weights}
meta.anlz.ad_cntrl$TE.random.abs <- abs( meta.anlz.ad_cntrl$TE.random )

#Assign Sig Or No
meta.anlz.ad_cntrl$Sig <- 'NO'
meta.anlz.ad_cntrl[meta.anlz.ad_cntrl$pval.random < 0.05, ]$Sig <- 'YES'

@@

#Annotate with Gene Names
library('biomaRt')
mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
genes <- as.character(Weights$ENSG[ !duplicated(Weights$ENSG) ])
G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","hgnc_symbol"),values=genes,mart= mart)

G_list$ensembl_gene_id <- as.character(G_list$ensembl_gene_id)
G_list$hgnc_symbol <- as.character(G_list$hgnc_symbol)

G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]
Miss <- cbind( ensembl_gene_id = as.character(Weights[ (Weights$ENSG %in% G_list$ensembl_gene_id) == F,]$ENSG),
               hgnc_symbol = rep("NA", length(as.character(Weights[ (Weights$ENSG %in% G_list$ensembl_gene_id) == F,]$ENSG)) )
               )
G_list <- rbind( Miss, G_list)
G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]

@@

## Store results in synapse
```{r syn.store, cache = FALSE, include=FALSE}
CODE <- syn_temp$store(synapseclient$Folder(name = "MetaAnalysis", parentId = parentId))
Syns_Used <- c( 'syn8456631', 'syn8484996', 'syn8466814', 'syn21292041', 'syn21285564', 'syn21291908' )

# Write results to files
fwrite(meta.anlz.ad_cntrl, file = 'meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
```

>>>>

```{r synapse.parameters, include=FALSE, cache=TRUE}
synapser::synLogin()
parentId = 'syn22351719';
activityName = 'Meta analysis of differential expression hg38 Reprocessed RNASeq';
activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data hg38 Reprocessed Data (6 brain regions - No CBE)';
thisFileName <- 'RNA-Seq_Meta_Analysis.Rmd'
# Github link
thisRepo <- getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

@@

```{r get.exp}
# Function to read tsv/csv files from synapse
downloadFile <- function(id, tissue){ 
  dat <- fread(synapser::synGet(id)$path, data.table = F)
  dat$tissue <- tissue
  return(data)
}
# Read covariates
covar.ids = c(ROSMAP_DLPFC = 'syn26340637 ', 
              ROSMAP_PCC = 'syn26350606',
              ROSMAP_ACC = 'syn26374583 ',
              MSSM_FP = 'syn26453689', 
              MSSM_IFG = 'syn26454154', 
              MSSM_PHG = 'syn26453943', 
              MSSM_STG = 'syn26454639', 
              MAYO_TCX = 'syn26374583')
#_tissue

all.used.ids = as.character(covar.ids)
covar =  covar.ids %>%
  lapply(downloadFile)
covar$ROSMAP_DLPFC$tissue <- 'DLPFC'
covar$ROSMAP_PCC$tissue <- 'PCC'
covar$ROSMAP_ACC$tissue <- 'ACC'
covar$MSSM_FP$tissue <- 'FP'
covar$MSSM_IFG$tissue <- 'IFG'
covar$MSSM_PHG$tissue <- 'PHG'
covar$MSSM_STG$tissue <- 'STG'
covar$MAYO_TCX$tissue <- 'TCX'

# Read adjusted expression
expr.ids = c(ROSMAP_DLPFC = 'syn26340641', 
              ROSMAP_PCC = 'syn26374587',
              ROSMAP_ACC = 'syn26350610',
              MSSM_FP = 'syn26453693', 
              MSSM_IFG = 'syn26454158', 
              MSSM_PHG = 'syn26453947', 
              MSSM_STG = 'syn26454643', 
              MAYO_TCX = 'syn26387667')

all.used.ids = c(all.used.ids, as.character(expr.ids))
downloadFile <- function(id, tissue){ 
  fread(synapser::synGet(id)$path, data.table = F)
}
expr =  expr.ids %>%
  lapply(downloadFile) %>%
  plyr::join_all(type = 'full')
```

```{r modify.covar}
covar = covar %>%
  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study')
```

## Perform fixed and random effect meta-analysis for each gene across 7 different brain regions
#### Meta-analysis of AD-CONTROL
```{r ad.cntrl}
# Mean and sd of gene expression in each tissue x diagnosis group
tissue.dx.summary = plyr::ddply(covar, .(tissue, diagnosis), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$feature,
             n = dim(x)[1],
             mn = rowMeans(y[,x$specimenID], na.rm = T),
             sd = apply(y[,x$specimenID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison - Exclude CBE
meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary[(tissue.dx.summary$tissue%in%'CBE')==F,], .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$tissue
  cntrl.effect = dplyr::filter(x, diagnosis == 'CT')
  rownames(cntrl.effect) = cntrl.effect$tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

@@

#### Adjust the AD-Control Model for weights
```{r weights}
meta.anlz.ad_cntrl <- meta.anlz.ad_cntrl[complete.cases(meta.anlz.ad_cntrl),]
meta.anlz.ad_cntrl$TE.random.abs <- abs( meta.anlz.ad_cntrl$TE.random )

#Assign Sig Or No
meta.anlz.ad_cntrl$Sig <- 'NO'
meta.anlz.ad_cntrl[meta.anlz.ad_cntrl$fdr.random < 0.05, ]$Sig <- 'YES'

@@

G_list <- fread(synapser::synGet('syn26453691')$path)
all.used.ids <- c(all.used.ids,'syn26453691')

G_list <- as.data.frame(G_list)
G_list$ensembl_gene_id <- as.character(G_list$ensembl_gene_id)
G_list$hgnc_symbol <- as.character(G_list$hgnc_symbol)

G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]

G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]

@@

## Store results in synapse
```{r syn.store, cache = FALSE, include=FALSE}
CODE <- synapser::synStore(synapser::Folder(name = "MetaAnalysis", parentId = parentId))
Syns_Used <- all.used.ids

# Write results to files
fwrite(meta.anlz.ad_cntrl, file = 'RNASeq_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  synapser::synStore( synapser::File( path='RNASeq_meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
fwrite(Weights, file = 'RNASeq_meta_weights.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  synapser::synStore( synapser::File( path='RNASeq_meta_weights.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis Weights', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
```
